|

### 2021-08-10 prova_knit_finale.Rmd
elimanato 1000000 particles da rimettere

\subsubsection{Convergence of the IS estimator}\label{pf_is_conv}
We can show that the auto-normalized IS estimator is both consistent and asymptotically normal.\
Consider the auto-normalized IS estimator $\hat \varphi= \sum_{n=1}^N\varphi (X^n)W^n$. We can write:
\begin{equation*}
   \hat \varphi_{AN} - \E_{\Q}(\varphi)=\E_{\Q^N}(\varphi)-\E_{\Q}(\varphi)=\frac{N^{-1}\sum_{n=1}^Nw(X^n)[\varphi(X^n)-\E_{\Q}(\varphi)]}{N^{-1}\sum_{n=1}^Nw(X^n)}
\end{equation*}
Let $\bar \varphi(X):=\varphi(X)-\E_{\Q}(\varphi)$. Applying the CLT to the numerator, assuming that $\E_{\M}(w^2\bar\varphi^2)<+\infty$, we get:
\begin{equation}
    \sqrt{N}\Bigg\{\frac{1}{N}\sum_{n=1}^Nw(X^n)\bar \varphi (X^n)\Bigg\}\implies \N \big(0,\E_{\M}(w^2\bar\varphi^2)\big)\label{is_conv_clt}
\end{equation}
where $\implies$ denotes convergence in distribution. Instead, for the denominator the strong LLN implies that:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n)\rightarrow^{\text{a.s.}} \E_{\M}(w) \label{is_conv_lln}
\end{equation}
Applying Slutsky's theorem, together with \eqref{is_conv_clt} and \eqref{is_conv_lln}, we obtain:
\begin{equation}
   \sqrt{N}\big[  \hat \varphi_{AN} - \E_{\Q}(\varphi)\big] \implies \N \Bigg(0,\frac{\E_{\M}(w^2\bar\varphi^2)}{[\E_{\M}(w)]^2}\Bigg)
\end{equation}
Therefore, the IS (auto-normalized) estimator for $\varphi$ is asymptotically normal. Consistency can be proved in an analogous way, applying the strong LLN and lemma \ref{lemma_imp_sam_1}:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n) \varphi (X^n)\rigtharrow^{\text{a.s.}}  \E_\M(\varphi \cdot w)=\E_\Q(\varphi)\cdot \E_\M(w)
\end{equation}
Together with \eqref{is_conv_lln}, we get strong convergence of the estimator (that implies consistency).
\begin{equation}
   \hat \varphi_{AN}=\sum_{n=1}^NW^n \varphi (X^n)\rigtharrow^{\text{a.s.}}  \E_\Q(\varphi)
\end{equation}

@@

```{r include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=5,nrow=5)
colnames(estimatedx)<-c("N","BPF","GPF", "APF","LWF")
estimatedx[,1]<-c("20","100","1000","10000","100000")

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
kbl(estimatedx, align="c", longtable = T, booktabs = T, caption="RMSE", digits=4)
```

>>>>

\subsubsection{Convergence of the IS estimator}\label{pf_is_conv}
We can show that the auto-normalized IS estimator is both consistent and asymptotically normal.\
Consider the auto-normalized IS estimator $\hat \varphi= \sum_{n=1}^N\varphi (X^n)W^n$. We can write:
\begin{equation*}
   \hat \varphi_{AN} - \E_{\Q}(\varphi)=\E_{\Q^N}(\varphi)-\E_{\Q}(\varphi)=\frac{N^{-1}\sum_{n=1}^Nw(X^n)[\varphi(X^n)-\E_{\Q}(\varphi)]}{N^{-1}\sum_{n=1}^Nw(X^n)}
\end{equation*}
Let $\bar \varphi(X):=\varphi(X)-\E_{\Q}(\varphi)$. Applying the CLT to the numerator, assuming that $\E_{\M}(w^2\bar\varphi^2)<+\infty$, we get:
\begin{equation}
    \sqrt{N}\Bigg\{\frac{1}{N}\sum_{n=1}^Nw(X^n)\bar \varphi (X^n)\Bigg\}\implies \N \big(0,\E_{\M}(w^2\bar\varphi^2)\big)\label{is_conv_clt}
\end{equation}
where $\implies$ denotes convergence in distribution. Instead, for the denominator the strong LLN implies that:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n)\xrightarrow{\text{a.s.}} \E_{\M}(w) \label{is_conv_lln}
\end{equation}
Applying Slutsky's theorem, together with \eqref{is_conv_clt} and \eqref{is_conv_lln}, we obtain:
\begin{equation}
   \sqrt{N}\big[  \hat \varphi_{AN} - \E_{\Q}(\varphi)\big] \implies \N \Bigg(0,\frac{\E_{\M}(w^2\bar\varphi^2)}{[\E_{\M}(w)]^2}\Bigg)
\end{equation}
Therefore, the IS (auto-normalized) estimator for $\varphi$ is asymptotically normal. Consistency can be proved in an analogous way, applying the strong LLN and lemma \ref{lemma_imp_sam_1}:
\begin{equation}
    \frac{1}{N}\sum_{n=1}^Nw(X^n) \varphi (X^n)\xrightarrow{\text{a.s.}}  \E_\M(\varphi \cdot w)=\E_\Q(\varphi)\cdot \E_\M(w)
\end{equation}
Together with \eqref{is_conv_lln}, we get strong convergence of the estimator (that implies consistency).
\begin{equation}
   \hat \varphi_{AN}=\sum_{n=1}^NW^n \varphi (X^n)\xrightarrow{\text{a.s.}}  \E_\Q(\varphi)
\end{equation}

@@

```{r eval=FALSE, include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=5,nrow=5)
colnames(estimatedx)<-c("N","BPF","GPF", "APF","LWF")
estimatedx[,1]<-c("20","100","1000","10000","100000")

@@

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results="asis"}
kbl(estimatedx, align="c", longtable = T, booktabs = T, caption="RMSE", digits=4)
```
